\section{Administration and Supervision}
\label{sec:AdminOperations}

This chapter covers day-to-day operational supervision: the admin UI, audit
logging, activity monitoring and session inspection.

\subsection{Admin UI}

The admin UI is a Svelte application served by Quarkus at
\texttt{/admin-ui/}. It is organized into seven routes:

\begin{description}
\item[Overview] Dashboard showing all known models with status summaries,
  head revisions, element/relationship/view counts and active session counts.
  Colour-coded status: green (no issues), yellow (stale), red (integrity
  problems or consistency check failures).

\item[Models] Detailed per-model view: catalog metadata, ACL configuration,
  head revision, estimated counts, tag list and available actions (create
  tag, rebuild, compact, export, delete). The model selector persists the
  chosen model in browser storage.

\item[Versions] Tag management: list existing tags with description,
  revision and creation timestamp. Create new tags from current HEAD. Export
  and import model packages.

\item[Access] Identity and authorization: view the current resolved identity,
  normalized roles, and per-model ACL configuration. Update ACLs for a
  model. Run authorization diagnostics.

\item[Sessions] Live session inspection: active WebSocket sessions per
  model, with user id, session id, connected-since timestamp and held locks.

\item[Operations] Recent model activity: operator-facing events emitted by
  the server for the selected model (join, disconnect, submit, accept,
  reject, lock, rebuild, compact, tag, export, import). Recent op batches
  with batch id, revision range, op count and timestamp.

\item[Audit] Admin and WebSocket audit trail: structured log events with
  resolved user id, action, request context and timestamp. Websocket
  lifecycle events (open rejected, join, message rejected, close).
\end{description}

The admin UI supports:

\begin{itemize}
\item Bootstrap auth inputs for \texttt{X-Archimesh-User} and
  \texttt{X-Archimesh-Roles}, plus an optional bearer token for OIDC mode.
\item Configurable poll interval (persisted in browser storage; set to 0 to
  disable auto-refresh).
\item When a bearer token is present, admin 401/403 failures explain likely
  token-expiry or missing-role/model-access problems.
\end{itemize}

\subsection{Audit logging}

The server emits two structured audit log streams:

\subsubsection{admin\_audit}

Emitted for admin diagnostic reads, model window reads and mutating admin
actions. Each event is an INFO-level log line with a stable JSON payload
containing:

\begin{itemize}
\item Resolved user id
\item Action name
\item Request context (model id, endpoint, parameters)
\item Timestamp
\end{itemize}

Rejected admin mutations (409, 400) are audited too.

Operational guidance:

\begin{itemize}
\item Ship these JSON lines to a dedicated audit sink or index if relied on
  for operator review.
\item Keep retention shorter than ordinary application logs if dashboard
  polling is frequent.
\item Avoid over-aggressive dashboard refresh intervals in production ---
  repeated window reads produce repeated audit events.
\item Treat \texttt{admin\_audit} as machine-readable output; parse the JSON
  payload, do not depend on free-form message text.
\end{itemize}

\subsubsection{ws\_audit}

Emitted for WebSocket lifecycle events:

\begin{itemize}
\item Rejected opens
\item Successful joins
\item Rejected WebSocket messages
\item Closes
\end{itemize}

Configurable via \texttt{ARCHIMESH\_AUDIT\_WEBSOCKET\_ACTIONS} (comma-separated
action names). Set to an empty value to disable WebSocket audit entirely.

Set \texttt{ARCHIMESH\_AUDIT\_WEBSOCKET\_VERBOSE=true} to also emit
\texttt{WebSocketMessageAccepted} events for accepted
\texttt{SubmitOps}, \texttt{AcquireLock}, \texttt{ReleaseLock} and
\texttt{Presence} messages.

\subsection{Supervision checklist}

Regular operational supervision tasks:

\begin{itemize}
\item Check the admin overview for models with stale head revisions (clients
  not submitting) or integrity issues.
\item Check the sessions page for unexpected disconnects or abandoned
  sessions.
\item Check the operations page for repeated \texttt{SubmitOpsRejected*}
  events --- these indicate client problems.
\item Check the audit page for unauthorized access attempts.
\item Monitor Neo4j disk usage --- op-log and materialized state grow with
  model activity.
\item Monitor Kafka topic sizes and consumer lag.
\item Periodically run \texttt{GET /admin/models/\{modelId\}/integrity} and
  check for orphan nodes or missing references.
\end{itemize}
