\section{Wire Protocol and Schema Reference}
\label{sec:AppendixSchemas}

This appendix documents the JSON schemas that define the Archimesh wire
protocol contract. The authoritative schema files live in
\filepath{schemas/}.

\subsection{Schema files}

\begin{longtable}{p{42mm}p{112mm}}
\toprule
File & Purpose \\
\midrule
\texttt{common.json} & Shared type definitions: \texttt{ModelId},
  \texttt{UserId}, \texttt{SessionId}, \texttt{EntityId},
  \texttt{RFC3339}, \texttt{Causal}, \texttt{Actor},
  \texttt{RevisionRange}, \texttt{JsonObject}. \\
\texttt{ops.json} & Operation type definitions for all supported op types:
  create, update, delete for elements, relationships, views, view objects,
  connections, folders; property and membership operations; plus strict
  notation JSON objects for view objects and connections. \\
\texttt{op-batch.json} & Op batch envelope: \texttt{SubmitOps} with
  \texttt{baseRevision}, \texttt{opBatchId}, \texttt{actor} and
  \texttt{ops} array. \\
\texttt{lock-event.json} & Lock event payload: action (acquired, released,
  expired), actor, targets, timestamp. \\
\texttt{presence.json} & Presence payload: modelId, userId, sessionId,
  viewId, selection array, cursor object. \\
\texttt{notation-contract.md} & Human-readable contract describing the
  notation field boundaries and parity requirements. \\
\bottomrule
\end{longtable}

\subsection{Message flow summary}

\begin{longtable}{p{30mm}p{30mm}p{94mm}}
\toprule
Direction & Type & Description \\
\midrule
Client $\to$ Server & \texttt{Join} & Request checkout. Optional
  \texttt{lastSeenRevision} and \texttt{ref}. \\
Client $\to$ Server & \texttt{SubmitOps} & Submit one op batch.
  \texttt{baseRevision}, \texttt{opBatchId}, \texttt{ops}. \\
Client $\to$ Server & \texttt{AcquireLock} & Acquire notation lock on
  targets. \texttt{ttlMs}. \\
Client $\to$ Server & \texttt{ReleaseLock} & Release notation lock on
  targets. \\
Client $\to$ Server & \texttt{Presence} & Broadcast cursor position and
  selection on a view. \\
\midrule
Server $\to$ Client & \texttt{CheckoutSnapshot} & Full materialized model
  state. \\
Server $\to$ Client & \texttt{CheckoutDelta} & Incremental ops since
  \texttt{lastSeenRevision}. \\
Server $\to$ Client & \texttt{OpsAccepted} & Batch commit confirmation with
  assigned revision range. \\
Server $\to$ Client & \texttt{OpsBroadcast} & Remote batch broadcast (Kafka
  fan-out or direct fallback). \\
Server $\to$ Client & \texttt{LockEvent} & Lock state change notification. \\
Server $\to$ Client & \texttt{PresenceBroadcast} & Remote cursor/selection
  update. \\
Server $\to$ Client & \texttt{Error} & Rejection or failure with
  \texttt{code} and \texttt{message}. \\
\bottomrule
\end{longtable}

\subsection{Common type definitions}

\subsubsection{Causal metadata}

\begin{lstlisting}[caption={Causal object}]
{
  "clientId": "<stable client or session identifier>",
  "lamport": <monotonic integer per client>,
  "opId": "<globally unique operation identity>"
}
\end{lstlisting}

\subsubsection{Actor}

\begin{lstlisting}[caption={Actor object}]
{
  "userId": "<user identifier>",
  "sessionId": "<session identifier>"
}
\end{lstlisting}

\subsubsection{Revision range}

\begin{lstlisting}[caption={Revision range object}]
{
  "from": <inclusive start revision>,
  "to": <inclusive end revision>
}
\end{lstlisting}

\subsection{Op batch envelope}

\begin{lstlisting}[caption={SubmitOps payload}]
{
  "type": "SubmitOps",
  "payload": {
    "baseRevision": <client revision>,
    "opBatchId": "<UUID>",
    "actor": { "userId": "...", "sessionId": "..." },
    "ops": [
      { "type": "CreateElement", ... },
      { "type": "UpdateViewObjectOpaque", ... }
    ]
  }
}
\end{lstlisting}

\subsection{Notation JSON objects}

View object and connection notation are validated against strict schemas
(\texttt{additionalProperties: false}):

\subsubsection{ViewObjectNotationJson}

Supported fields: \texttt{x}, \texttt{y}, \texttt{width}, \texttt{height},
\texttt{type}, \texttt{alpha}, \texttt{lineAlpha}, \texttt{lineWidth},
\texttt{lineStyle}, \texttt{textAlignment}, \texttt{textPosition},
\texttt{gradient}, \texttt{iconVisibleState},
\texttt{deriveElementLineColor}, \texttt{fillColor}, \texttt{lineColor},
\texttt{font}, \texttt{fontColor}, \texttt{iconColor},
\texttt{imagePath}, \texttt{imagePosition}, \texttt{name},
\texttt{documentation}.

\subsubsection{ConnectionNotationJson}

Supported fields: \texttt{type}, \texttt{nameVisible},
\texttt{textAlignment}, \texttt{textPosition}, \texttt{lineWidth},
\texttt{name}, \texttt{lineColor}, \texttt{font}, \texttt{fontColor},
\texttt{documentation}, \texttt{bendpoints}.

\subsection{Error codes}

\begin{longtable}{p{46mm}p{108mm}}
\toprule
Code & Meaning \\
\midrule
\texttt{REVISION\_AHEAD} & Client \texttt{baseRevision} exceeds server
  head. Client must rejoin. \\
\texttt{PRECONDITION\_FAILED} & Referenced entity does not exist, or
  notation contains unknown keys. \\
\texttt{LOCK\_CONFLICT} & Another client holds a conflicting notation lock. \\
\texttt{MODEL\_NOT\_FOUND} & The requested model is not registered. \\
\texttt{REF\_NOT\_FOUND} & The requested tag name does not exist. \\
\texttt{REF\_READ\_ONLY} & Write attempted on a tagged (read-only) ref. \\
\bottomrule
\end{longtable}

\subsection{Export/import format}

The admin export format is a JSON package containing:

\begin{lstlisting}[caption={Export package structure}]
{
  "format": "archi-model-export-v1",
  "exportedAt": "<UTC timestamp>",
  "model": { "modelId": "...", "modelName": "...", "headRevision": 42 },
  "accessControl": { "configured": true, "adminUsers": [...], ... },
  "snapshot": { "format": "archimate-materialized-v1", ... },
  "opBatches": [ { "modelId": "...", "opBatchId": "...", ... } ],
  "tags": [ { "tagName": "...", "revision": 24, "snapshot": {...} } ]
}
\end{lstlisting}

The package format is documented in full in \filepath{server/EXPORT\_FORMAT.md}.
