\section{Messaging}
\label{sec:Messaging}

Kafka is used as the ordered fan-out path for accepted Archimesh operation
batches, lock events and presence events.

\subsection{Topic layout}

Three topic kinds exist per model:

\begin{longtable}{p{46mm}p{54mm}p{54mm}}
\toprule
Topic pattern & Content & Partitioning \\
\midrule
\texttt{archimesh.model.<modelId>.ops} & Committed op batches &
  Exactly 1 partition (required) \\
\texttt{archimesh.model.<modelId>.locks} & Lock acquire/release/expire &
  1 partition \\
\texttt{archimesh.model.<modelId>.presence} & Cursor/selection broadcasts &
  1 partition (or keyed by sessionId) \\
\bottomrule
\end{longtable}

The topic prefix is configurable via \texttt{KAFKA\_TOPIC\_PREFIX} (default
\texttt{archimesh.model}).

\subsection{Single-partition requirement}

Each model's ops topic \textbf{must} have exactly one partition. This is a
hard constraint, not a recommendation. Multiple partitions would allow
operations to be consumed in different orders by different clients, breaking
the total order guarantee that the deterministic merge semantics depend on.

Topic keys:

\begin{itemize}
\item ops: constant key = \texttt{<modelId>}
\item locks: constant key = \texttt{<modelId>}
\item presence: key = \texttt{<sessionId>} (or \texttt{<userId>})
\end{itemize}

\subsection{Producer}

The \texttt{KafkaPublisherImpl} uses a \texttt{KafkaProducer<String,{}String>}
with:

\begin{itemize}
\item \texttt{acks=all} --- wait for all in-sync replicas
\item \texttt{retries=3}
\item \texttt{delivery.timeout.ms=120000}
\item String serializers for both key and value
\end{itemize}

Publishing is asynchronous. The \texttt{publishOps} method returns a
\texttt{CompletableFuture<Void>} that completes when the broker acknowledges
the send. If the send fails, the server falls back to broadcasting
\texttt{OpsBroadcast} directly to connected WebSocket sessions.

\subsection{Consumer}

The \texttt{KafkaConsumerImpl} runs a single-threaded consume loop with:

\begin{itemize}
\item Group id: \texttt{archimesh-server} (configurable via
  \texttt{KAFKA\_CONSUMER\_GROUP\_ID})
\item Auto offset reset: \texttt{earliest}
\item Topic subscription via regex pattern matching all
  \texttt{<prefix>.<modelId>.<kind>} topics
\item 500 ms poll timeout
\end{itemize}

On receiving a record, the consumer parses the topic name to extract the
model id and kind, then broadcasts the payload to all sessions registered
for that model:

\begin{description}
\item[ops] Broadcasts \texttt{OpsBroadcast} to all sessions for the model.
  The submitter suppresses its own echo.
\item[locks] Broadcasts \texttt{LockEvent}.
\item[presence] Broadcasts \texttt{PresenceBroadcast}.
\end{description}

The consumer is resilient to Kafka unavailability: on connection failure it
sleeps and retries. While the consumer is down, the direct broadcast fallback
in the submit path (Section~\ref{sec:ServerApi}) ensures connected clients
still receive updates.

\subsection{Retention}

\begin{description}
\item[ops] Long retention if Kafka is used for rebuild, otherwise moderate.
  The Neo4j op-log is the durable log; Kafka is the fan-out path.
\item[locks] Short retention (minutes). Lock state is ephemeral.
\item[presence] Very short retention (minutes). Presence is transient.
\end{description}

\subsection{Fallback broadcast}

If Kafka publish fails (broker unavailable, topic error, serialization
failure), the \texttt{ArchimeshService} catches the exception and broadcasts
\texttt{OpsBroadcast} directly to all sessions registered for the model. This
ensures that connected clients on the same server instance still receive
remote updates even when the Kafka fan-out path is unavailable.

The fallback uses the same \texttt{OpsBroadcastMessage} envelope as the
consumer path, so clients process it identically. Echo suppression on the
submitting client prevents double-application.

\subsection{Configuration}

\begin{longtable}{p{50mm}p{54mm}p{50mm}}
\toprule
Property & Env var & Default \\
\midrule
\texttt{kafka.bootstrap.servers} & \texttt{KAFKA\_BOOTSTRAP\_SERVERS} &
  \texttt{localhost:9092} \\
\texttt{kafka.topic.prefix} & \texttt{KAFKA\_TOPIC\_PREFIX} &
  \texttt{archimesh.model} \\
\texttt{kafka.consumer.group-id} & \texttt{KAFKA\_CONSUMER\_GROUP\_ID} &
  \texttt{archimesh-server} \\
\texttt{kafka.consumer.auto-offset-reset} &
  \texttt{KAFKA\_CONSUMER\_AUTO\_OFFSET\_RESET} &
  \texttt{earliest} \\
\bottomrule
\end{longtable}
